\chapter*{Acknowledgments}

This book has been the result of many years of experimentation in my own life.
I would like to thank my family first, who has had to put up with all of my 
experiments in living in new and often strange ways.  I thank especially my
wife, who still loved me even after I converted the children's playset into 
a chicken coop, and whose only warning that we were getting these new members
of our family was a text message from me saying,
``I took the boys cruising to pick up some chicks.''  

I want to extend special thanks to the people who have contributed directly
to this book.  I had an outstanding team to help put this together, including
Robert Lamar, a fantastic book designer, Meghan Barnes, an excellent editor,
and several preview readers who gave great feedback.

I also want to thank the people I work with every day.  As a software engineer
by trade, I am almost always focused on the practical and the calculable.
Working with people whose focus is on the aesthetic has helped me to 
view the world in wider terms, and to think about more than just the numbers.

I also want to thank my professors from Phillips Theological Seminary.
While I can't overemphasize how much I disagree with their overall view of life
and theology, what I learned there was how to apply theological to
practical situations, and I hope this book reflects well on this training.  

Finally, I want to thank the Church---not just where I attend now, but the 
whole body of believers.  I have travelled through many hard times, and the
Church in its many forms has always been there to help me.

There are many other people to thank, and one of the problems with acknowledgments
is that if one were to start listing the people who have influenced you, helped you,
or contributed to your growth as a person, you might not have room for the rest
of the book.  I just want you to know that I love and appreciate you all!
